Succinct representations for abstract interpretation

### 2011-11-01 papers/PLDI2012/implicitization.tex
X^d et X^s
typos + retrait TODO

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i^s$ the abstract value
associated to $p_i$ :
$$f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge x \in X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
x \notin X_j^s)$$

@@

Also, to prevent the algorithm from adding too many paths at a time,
and in particular to prevent the algorithm from trying to re-add the
same path infinitely many times, we use another abstract value
associated to a control point $p_j$, noted $X_j^d$, which is distinct
from $X_j^s$, and initialized to $X_j^s$. This value is updated when a
point $p_j$ is the target of a new path, and is used instead of
$X_j^s$ in the SMT formula. This way, further SMT queries do not
compute other paths with the same source and destination if it is not
needed (because these new paths would not make $X_j^d$ grow, hence
would not be returned by the SMT solver).

@@

For each control point $p_i$, we compute a disjunctive invariant
$\bigvee_{1\leq j \leq m_i} X_{i,j}$. We denote by $n_i$ the number of
distinct paths starting at $p_i$.
To perform the analysis, one chooses an integer $\delta_i \in [1,m_i]$, and
a mapping function $\sigma_i: [1,m_i] \times [1,n_i] \mapsto [1,m_i]$.
\MM{C'est moi ou le $\tau_{i,k}$ ne correspond pas exactement à la
  notation du début ? Au début, $\tau_{i,k}$ est la transition de $i$
  à $k$, alors que là, on est dans le multigraphe donc il y a
  plusieurs arrêtes entre deux points. Si je comprends bien,
  $\tau_{i,k}$ est le $k$-ième chemin partant de $i$, c'est ça ?
  Je mettrais bien ``... and for the $k$-th path starting from $i$,
  denoted $\tau_{i,k}$, the image...'' si c'est bien ça. D'ailleurs,
  faut-il forcément parler de boucle ici ? La technique ne s'applique
  pas à des chemins quelconques ?}
  \JH{J'ai corrige\dots}
  The $k$-th path starting fom $p_i$ is denoted $\tau_{i,k}$.
  The image of the $j$-th disjunct $X_{i,j}$ by the path $\tau_{i,k}$ is then
  joined with $X_{i,\sigma(j,k)}$.
Initially, the $\delta_i$-th abstract value contains the initial states of
$p_i$, and all other abstract values contain~$\emptyset$.

@@

\begin{eqnarray*}
g(p_i) = & \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg b_j^s  \\
 & \wedge 
\displaystyle\bigvee_{1 \leq k \leq m} (d_k \wedge x \in X_{i,k} \wedge \bigwedge_{l \neq k}
\neg d_l) \\
 & \wedge
\displaystyle\bigvee_{j \in Succ(i)} 
(b_j^d \wedge \bigwedge_{1 \leq k \leq m} (x \notin X_{j,k}))
\end{eqnarray*}

@@

The analysis of linear equalities (PKEQ) performs very fast compared
to other abstract domains, but yields very imprecise invariants:
it only detects relations of the form $\sum_i a_i x_i = C$ where $a_i$ and $C$ are constants.
The comparison between the domain of convex polyhedra and linear congruences
shows the percentage of points where the use of congruences refines the
invariant. This is the case for $1.04\%$ of the control points in $P_R$.
\DM{Je croyais qu'on ne pouvait pas utiliser PPLGRID?}
\JH{On peut. C'est PKGRID qui ne fonctionne pas}

>>>>

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i^s$ the abstract value
associated to $p_i$ :
$$f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
 \neg X_j^s)$$

@@

Also, to prevent the algorithm from adding too many paths at a time,
and in particular to prevent the algorithm from trying to re-add the
same path infinitely many times, we use another abstract value
associated to a control point $p_j$, noted $X_j^d$, which is distinct
from $X_j^s$, and initialized to $X_j^s$. This value is updated when a
point $p_j$ is the target of a new path, and is used instead of
$X_j^s$ in the SMT formula. We thus check the satisfiability of the formula
$f'(p_i)$, where:
$$f'(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
\neg X_j^d)$$

This way, further SMT queries do not
compute other paths with the same source and destination if it is not
needed (because these new paths would not make $X_j^d$ grow, hence
would not be returned by the SMT solver).

@@

For each control point $p_i$, we compute a disjunctive invariant
$\bigvee_{1\leq j \leq m_i} X_{i,j}$. We denote by $n_i$ the number of
distinct paths starting at $p_i$.
To perform the analysis, one chooses an integer $\delta_i \in [1,m_i]$, and
a mapping function $\sigma_i: [1,m_i] \times [1,n_i] \mapsto [1,m_i]$.
  The $k$-th path starting fom $p_i$ is denoted $\tau_{i,k}$.
  The image of the $j$-th disjunct $X_{i,j}$ by the path $\tau_{i,k}$ is then
  joined with $X_{i,\sigma(j,k)}$.
Initially, the $\delta_i$-th abstract value contains the initial states of
$p_i$, and all other abstract values contain~$\emptyset$.

@@

\begin{eqnarray*}
g(p_i) = & \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg b_j^s  \\
 & \wedge 
\displaystyle\bigvee_{1 \leq k \leq m_i} (d_k \wedge X_{i,k} \wedge \bigwedge_{l \neq k}
\neg d_l) \\
 & \wedge
\displaystyle\bigvee_{j \in Succ(i)} 
(b_j^d \wedge \bigwedge_{1 \leq k \leq m_i} (\neg X_{j,k}))
\end{eqnarray*}

@@

The analysis of linear equalities (PKEQ) performs very fast compared
to other abstract domains, but yields very imprecise invariants:
it only detects relations of the form $\sum_i a_i x_i = C$ where $a_i$ and $C$ are constants.
The comparison between the domain of convex polyhedra and linear congruences
shows the percentage of points where the use of congruences refines the
invariant. This is the case for $1.04\%$ of the control points in $P_R$.
